---
title: "Exploratory analysis"
author: "Shumei Liu"
date: "2024-10-08"
output: github_document
---

```{r, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
weather_df = read_csv("data/weather_df.csv") |>
  mutate(month = lubridate::floor_date(date, unit = "month"))

weather_df
```

## Initial numeric explorations

```{r}
weather_df |> 
  ggplot(aes(x = prcp)) + 
  geom_histogram()
```

```{r}
weather_df |> 
  filter(prcp >= 1000)
```

```{r}
weather_df |> 
  filter(tmax >= 20, tmax <= 30) |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) + 
  geom_point(alpha = .75)
```

## `group_by`

```{r}
weather_df |>
  group_by(name, month) |>
  ungroup()
```

## counting things

count month observations

```{r}
weather_df |>
  group_by(month) |>
  summarize(n_obs = n()) # n of each month
```

```{r}
weather_df |>
  group_by(name, month) |>
  summarize(n_obs = n())
```

we can use `count()`

```{r}
weather_df |>
  count(month, name = "n_obs")
```

**NEVER** use base R's `table`

```{r, eval=FALSE}
weather_df |>
  pull(month) |> 
  table() #do NOT use this function
```

other helpful counters

```{r}
weather_df |>
  group_by(month) |>
  summarize(
    n_obs = n(),
    n_days = n_distinct(date)) #每月不同的每一天
```

## A digression on 2x2 tables

```{r}
weather_df |> 
  drop_na(tmax) |> 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ "" #if missing anything
  )) |> 
  filter(name != "Molokai_HI") |> 
  group_by(name, cold) |> 
  summarize(count = n())
```


```{r}
weather_df |> 
  drop_na(tmax) |> 
  mutate(cold = case_when(
    tmax <  5 ~ "cold",
    tmax >= 5 ~ "not_cold",
    TRUE     ~ ""
  )) |> 
  filter(name != "Molokai_HI") |> 
  janitor::tabyl(name, cold) #2x2 table
```

## General summaries

You can do lots of summaries.

```{r}
weather_df |>
  group_by(month) |>
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE), #do NOT conclude NA value when doing calculation
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax))
```

You can group by more than one variable.

```{r}
weather_df |>
  group_by(name, month) |>
  summarize(
    mean_tmax = mean(tmax),
    median_tmax = median(tmax))
```

This is a dataframe! Se we can make a plot.

```{r}
weather_df |>
  group_by(name, month) |>
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) |>
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
    geom_point() + geom_line() + 
    theme(legend.position = "bottom")
```

Suppose you want to summarize many columns.

If you want to summarize multiple columns using the same summary, the `across` function is helpful.

```{r}
weather_df |>
  group_by(name, month) |>
  summarize(across(tmin:prcp, mean))
```

Reminder: sometimes your results are easier to read in another format.

```{r}
weather_df |>
  group_by(name, month) |>
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) |> 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) |> 
  knitr::kable(digits = 1) #format results in an easier way to read, create a table
```

## `group_by` and `mutate`

```{r}
weather_df |>
  group_by(name) |>
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax) |> 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 
```

## Window functions

what about window functions?

ranking

```{r}
weather_df |>
  group_by(name, month) |>
  mutate(temp_ranking = min_rank(tmax))
```

```{r}
weather_df |>
  group_by(name, month) |>
  filter(min_rank(tmax) < 2)
```

keep the three days with the highest max temperature

```{r}
weather_df |>
  group_by(name, month) |>
  filter(min_rank(desc(tmax)) < 4)
```

lag 滞后

```{r}
weather_df |>
  group_by(name) |>
  mutate(lag_temp = lag(tmax)) 
#创建新的一列并向下移动一格，例如此处lag(tmax)，所以复制tmax重命名为lag_temp，并且相应的数据向下移动一格
```

```{r}
weather_df |>
  group_by(name) |>
  mutate(temp_change = tmax - lag(tmax))
```

```{r}
weather_df |>
  group_by(name) |>
  mutate(temp_change = tmax - lag(tmax)) |>
  summarize(
    temp_change_sd = sd(temp_change, na.rm = TRUE),
    temp_change_max = max(temp_change, na.rm = TRUE))
```

## Quick notes

summarize only gets you so far.

## example 1

Learning Assessment: In the PULSE data, the primary outcome is BDI score; it’s observed over follow-up visits, and we might ask if the typical BDI score values are roughly similar at each. Try to write a code chunk that imports, cleans, and summarizes the PULSE data to examine the mean and median at each visit. Export the results of this in a reader-friendly format.

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") |>
  select(id, visit, everything()) |>
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) |>
  arrange(id, visit)

pulse_data |> 
  group_by(visit) |> 
  summarize(
    mean_bdi = mean(bdi, na.rm = TRUE),
    median_bdi = median(bdi, na.rm = TRUE)) |> 
  knitr::kable(digits = 3)
```

This quick summary suggests a relatively large drop in the typical BDI score from baseline to 1 month, with small or no changes thereafter.

## example 2

Learning Assessment: In the FAS data, there are several outcomes of interest; for now, focus on post-natal day on which a pup is able to pivot. Two predictors of interest are the dose level and the day of treatment. Produce a reader-friendly table that quantifies the possible associations between dose, day of treatment, and the ability to pivot.

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv") |>
  janitor::clean_names() |>
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv") |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data |> 
  group_by(dose, day_of_tx) |> 
  drop_na(dose) |> 
  summarize(mean_pivot = mean(pd_pivot, na.rm = TRUE)) |> 
  pivot_wider(
    names_from = dose, 
    values_from = mean_pivot) |> 
  knitr::kable(digits = 3)
```

These results may suggest that pups in the control group are able to pivot earlier than pups in the low-dose group, but it is unclear if there are differences between the control and moderate-dose groups or if day of treatment is an important predictor.
