---
title: "Visualization"
subtitle: "Part I"
author: "Shumei Liu"
date: "2024-10-08"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges) #run the ridge plot
```

## Load the weather data

```{r}
weather_df = read_csv("data/weather_df.csv")

weather_df
```

The following code does not work! Don't know why!

`weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())`


## Scatterplots!

Create my first scatterplot ever.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

New approach, same plot.

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Save and edit a plot object.

```{r}
weather_plot =
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()
```

## Advanced scatterplot ...

Start with the same one and make it fancy!

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, colour = name)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

What about the `aes` placement?

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(colour = name)) +
  geom_smooth() #only have a blue line, because the color only use for the geom_point
```

Let's facet some things!

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, alpha = tmin, colour = name)) + #"alpha" is 点的透明度
  geom_point() +
  geom_smooth(se = FALSE, size = 1) +
  facet_grid(. ~ name) #"." means the row does not need to be defined, only use "name" to define the columns
```

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, colour = name)) +
  geom_point(alpha = .3, size = .8) +
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .) #create the row not the column
```

Let's combine some elements and try a new plot.

```{r}
weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) + #"size = prcp" 降雨量
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```


## Some small notes

How many geoms have to exist?

You can have whatever geoms you want. 

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

You can use a neat geom!

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_bin2d()
```

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_density_2d() +
  geom_point(alpha = .3)
```


## Univariate plots 单变量

Histograms are really great.

```{r}
weather_df %>%
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Can we add color?

```{r}
weather_df %>%
  ggplot(aes(x = tmin, fill = name)) + #"fill" 填充颜色
  geom_histogram(position = "dodge") #让不同颜色的直方图并列而不重合
```

```{r}
weather_df %>%
  ggplot(aes(x = tmin)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

Let's try a new geometry!

```{r}
weather_df %>%
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3) 
```

```{r}
weather_df %>%
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3, adjust = .5) #"adjust"看到单个数据的小突起
```

What about box plot??

```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```

Trendy plots :-)

```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = .5) +
  stat_summary(fun = "median")
```

similar

```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = .5) +
  stat_summary()
```

Ridge plots -- the most popular plot of 2017 山脊图
`library(ggridges)`

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

## Save and Embed 嵌入

Let's save a scatterplot.

```{r}
weather_plot =
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)

ggsave("results/weather_plot.pdf", weather_plot, width = 8, height = 5)
```

What about embedding?

```{r, fig.width=6, fig.height=3}
weather_plot
```

Embed at different size.

```{r, fig.width=12, fig.height=6}
weather_plot
```

