---
title: "Visualization"
subtitle: "Part I"
author: "Shumei Liu"
date: "2024-10-08"
output: github_document
---

```{r}
library(tidyverse)
library(patchwork)
```

## Load the weather data

```{r}
weather_df = read_csv("data/weather_df.csv")

weather_df
```

## Remember this plot?

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
```

## Labels

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location", #图例的名称
    caption = "Data from the rnoaa package"
  )
```

## Scales

Start with the same plot

only scale x aixs
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15º C", "0", "15º C"))
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ºC", "0", "15"),
    limits = c(-20, 30)) + 
  scale_y_continuous(
    trans = "sqrt", #also can use `log`...
    position = "right")
```

Look at color scales

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  scale_color_hue(h = c(100, 300)) #`scale`尺度, `hue`色调
```

Trying to create your own color scheme usually doesn’t go well, so use the `viridis` package instead.

```{r}
ggp_temp_plot = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis( 
    name = "Location", 
    discrete = TRUE
  ) #in viridis color绿藻色, `scale_color_viridis` only be used in continuous situation, so we need `discrete = TRUE`

ggp_temp_plot
```

## Themes

Shift the legend.

```{r}
ggp_temp_plot + 
  theme(legend.position = "bottom")
```

Change the overall theme.

```{r}
ggp_temp_plot + 
  theme_bw() + #theme black and white
  theme(legend.position = "bottom")
```

```{r}
#use this one the best for professor
ggp_temp_plot + 
  theme_minimal() + 
  theme(legend.position = "bottom")
```

```{r}
ggp_temp_plot + 
  theme_classic() + 
  theme(legend.position = "bottom")
```

```{r}
ggp_temp_plot + 
  ggthemes::theme_excel() + 
  theme(legend.position = "bottom")
```

```{r}
ggp_temp_plot + 
  ggthemes::theme_economist() + 
  theme(legend.position = "bottom")
```

Revisit the plot showing `tmax` against `date` for each location. Use labels, scale options, and theme changes to improve the readability of this plot.

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) + 
  geom_point(aes(size = prcp), alpha = .75) + 
  labs(
    title = "Temperature plot",
    x = "Date",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(discrete = TRUE) + 
  theme_minimal() + 
  theme(legend.position = "bottom")
```


## Setting options

A few other figure preferences in global options declared at the outset of each .Rmd file, this code chunk just gets copy-and-pasted to the beginning of every new file.

```{r, eval=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
确保全局的plot都保持以上相同属性

## Data argument in `geom_*`

```{r}
central_park_df = 
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df = 
  weather_df |> 
  filter(name == "Molokai_HI")

ggplot(data = molokai_df, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park_df) #one dataset use point plot, the other one use line plot
```





